---
title: "officer R package"
output:
  github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

The officer package lets R users manipulate Word (`.docx`) and 
PowerPoint (`*.pptx`) documents. In short, one can add images, 
tables and text into documents from R. An initial document can 
be provided; contents, styles and properties of the original 
document will then be available.

*This package is close to ReporteRs as it produces Word and PowerPoint files but it is faster, do not require `rJava` (but `xml2`) and has less functions that will make it easier to maintain.*

```{r results='asis', echo=FALSE}
logo_file <- tempfile(fileext = ".png")
rsvg::rsvg_png("/Users/davidgohel/Github/officer/inst/officerlogo.svg", file = logo_file, 
               width = 181, height = 209)

htmltools::span( htmltools::HTML(sprintf('<img src="data:image/png;base64,%s"/>',
        gdtools:::base64_file_encode(filename = logo_file) )))
```

> Make corporate reporting with minimum hassle

[![Travis-CI Build Status](https://travis-ci.org/davidgohel/officer.svg?branch=master)](https://travis-ci.org/davidgohel/officer)
[![Build status](https://ci.appveyor.com/api/projects/status/github/davidgohel/officer?branch=master)](https://ci.appveyor.com/project/davidgohel/officer/branch/master)
[![version](http://www.r-pkg.org/badges/version/officer)](https://CRAN.R-project.org/package=officer)
![cranlogs](http://cranlogs.r-pkg.org./badges/officer)
![Active](http://www.repostatus.org/badges/latest/active.svg)


## Word documents

The `read_docx()` function will read an initial Word document (an 
empty one by default) and let you modify its content later.

The package provides functions to add R outputs into a Word document:

* images: produce your plot in png or emf files and add them into the document, 
  as a whole paragraph or inside a paragraph.
* tables: add data.frames as tables, format is defined by the associated Word table style.
* text: add text as paragraphs or inside an existing paragraph, format is defined by the associated Word paragraph and text styles.
* field codes: add Word field codes inside paragraphs. Field codes is an old 
  feature of MS Word to create calculated elements such as tables of contents, 
  automatic numbering and hyperlinks.

In a Word document, one can use cursor functions to reach the
beginning or end of a document, or a particular paragraph containing a
given text. This *cursor* concept has been implemented to make the
post processing of files easier.

File generation is performed with the `print` function.

### import Word document in a data.frame

The function `docx_summary()` reads and imports content of a Word 
document into a tibble object. The function handles paragraphs, 
tables and section breaks.


## PowerPoint documents


The function `read_pptx()` will read an initial PowerPoint document 
(an empty one by default) and let you modify its content later.

The package provides functions to add R outputs into existing or new PowerPoint slides:

* images: produce your plot in png or emf files and add them in a slide.
* tables: add data.frames as tables, format is defined by the associated PowerPoint table style.
* text: add text as paragraphs or inside an existing paragraph, format is defined in the corresponding layout of the slide.

In a PowerPoint document, one can set a slide as selected and reach a particular shape (and remove it or add text). 

ile generation is performed with the `print()` function.

### import PowerPoint document in a data.frame

The `pptx_summary()` function reads and imports content of a PowerPoint 
document into a tibble object. The function handles paragraphs, tables and images.

## Extensions

### Tables and package `flextable`

The package [flextable](https://github.com/davidgohel/flextable) brings a full API to produce nice tables and use them with `officer`. Tables can be written in 
PowerPoint documents and Word documents. An option is available to render 
flextables in rmarkdown (HTML and Word outputs).

### Vector graphics with package `rvg`

The package [rvg](https://github.com/davidgohel/rvg) brings an API to produce nice vector graphics that can be embedded in PowerPoint documents or Excel workbooks with `officer`.

### Native office charts with package `mschart`

The package [mschart](https://github.com/ardata-fr/mschart) combined with `officer` can 
produce native office charts in PowerPoint and Word documents.


## Installation

You can get the development version from GitHub:

```{r eval = FALSE}
devtools::install_github("davidgohel/officer")
```

Or the latest version on CRAN:

```{r eval = FALSE}
install.packages("officer")
```

## Ressources

### Online documentation

The help pages are located at https://davidgohel.github.io/officer/.

### Getting help

This project is developped and maintained on my own time. In order to help me to maintain the 
package, do not send me private emails if you only have questions about how to use the package.
Instead, visit Stackoverflow, `officer` has its own tag [Stackoverflow link](http://stackoverflow.com/questions/tagged/officer)! 
I usually read them and answer when possible. 

```{r echo=FALSE, results='asis'}
cat(paste( readLines("CONTRIBUTING.md"), collapse = "\n"))
```

