---
title: "pptx device"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pptx device}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> Package `officer` provides a set of functions to manipulate PowerPoint presentations from R.

*Package `magrittr` can be used with `officer` as it will make the workflow easier to read.*

```{r}
library(officer)
library(magrittr)
```

## Introduction

Use the function `read_pptx` to create an r object representing a PowerPoint document.
It takes one argument `path`: a initial PowerPoint file. If none is provided, this file will be an empty document
located in the package directory. Formats and available slide layouts will be those available in the template file.

### Create a PowerPoint device

```{r}
my_pres <- read_pptx() 
```

### Add a slide

Next thing to do will most often be to add a new slide, this can be achieved with function `add_slide`. It requires 3 arguments: 

- a pptx object
- a slide layout name
- a master layout name

```{r}
my_pres <- my_pres %>% 
  add_slide(layout = "Two Content", master = "Office Theme")
```

#### master layouts and slide layouts

`officer` uses a PowerPoint file as initial document. This is the original
PowerPoint document where all slides layouts, shapes (placeholders) and
styles come from. Major points to be aware of are:

* Slide layouts are relative to a master layout. A document can contain more than one 
master layout ; a master layout can contain more than one slide layout. 
* A slide layout inherits design properties from its master layout file but 
some properties can be overwritten.
* Designs and formating properties of layouts and shapes (placeholders 
in a layout) are defined within the initial document. There is no R function 
to modify these values, they must be defined in the initial document.

**`layout` and `master` values must match values from the initial document.** 
Layout names and master layout names are not available in a tidy view within PowerPoint, but theses can be read easily with function `layout_summary`.

```{r}
layout_summary(my_pres)
```

### Add text content into a placeholder

Use function `ph_with_text` to add text into a shape defined in the slide layout associated to the current slide.

```{r}
my_pres <- my_pres %>% 
  ph_with_text(type = "title", str = "A title") %>%
  ph_with_text(type = "ftr", str = "A footer") %>%
  ph_with_text(type = "dt", str = format(Sys.Date())) %>%
  ph_with_text(str = "text 1", type = "body", index = 1) %>%
  ph_with_text(str = "text 2", type = "body", index = 2)
```

#### Available shapes per slide layout

Function `layout_properties` provides details about available shapes of a slide layout:

```{r}
layout_properties ( x = my_pres, layout = "Two Content", master = "Office Theme" ) %>% head()
```

### Write the PowerPoint

File can be generated using function `print` and argument `target`:

```{r}
print(my_pres, target = "example.pptx" )
```

## Slides manipulation

There are 3 functions to let you manipulate slides: add_slide, remove_slide and on_slide.

### Add a slide 

A slide can be added with function `add_slide`.

```{r}
my_pres <- read_pptx() %>% 
  add_slide(layout = "Two Content", master = "Office Theme") %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  add_slide(layout = "Title Only", master = "Office Theme")
length(my_pres)
```

### Remove a slide 

A Slide can be removed with function `remove_slide`.

```{r}
my_pres <- my_pres %>% remove_slide(index = 1)
length(my_pres)
```

### Select current slide 

Selected slide can be set with function `on_slide`. 

```{r}
my_pres <- my_pres %>% on_slide(index = 1)
```

## Add content into a placeholder

### Add text

Use function `ph_with_text` to add text into a placeholder (or a shape). 
Argument `type` specify the placeholder of the associated layout to be 
added as new shape (`index` is to be used when an `type` is not unique 
in the slide layout).

```{r}
doc <- read_pptx() 
doc <- doc %>%
  add_slide(layout = "Two Content", master = "Office Theme") %>%
  ph_with_text(type = "body", str = "A first text", index = 1) %>%
  ph_with_text(type = "body", str = "A second text", index = 2) %>%
  ph_with_text(type = "title", str = "A title") %>%
  ph_with_text(type = "ftr", str = "Slide footer") %>%
  ph_with_text(type = "dt", str = format(Sys.Date()))
```

Use `layout_properties` to see what are available placeholders in the slide layout. 

### Add image

Use function `ph_with_img` to add an image into a placeholder. Again, 
argument `type` specify the placeholder of the associated layout to be added 
as new shape (`index` is to be used when an `type` is not unique in the slide layout).

```{r}
doc <- read_pptx() 
doc <- doc %>%
  add_slide(layout = "Two Content", master = "Office Theme") %>%
  ph_with_text(type = "body", str = "body (index 1) is text", index = 1)

img.file <- file.path( Sys.getenv("R_HOME"), "doc", "html", "logo.jpg" )
f_exists <- file.exists(img.file)
if( f_exists ){
  doc <- ph_with_img(x = doc, type = "body", index = 2, src = img.file, height = 1.06, width = 1.39 )
}
```


## Remove content from a slide

Use `slide_summary` to easily identify shapes in the slide that can be removed.

```{r}
slide_summary(doc)
```

In the following example, shape corresponding to `type "body"` and `id "3"` will be removed from the 
current slide:

```{r}
doc <- ph_remove(x = doc, type = "body", id_chr = "3")
```


## Append text sequentially in a shape

### Add to an empty new placeholder

`ph_empty` will add a new empty placeholder in the current slide. 

```{r}
my_pres <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_empty(type = "body")
```

As there is not paragraph yet in the new shape, function `ph_add_par` will be used to add a new paragraph. Then `ph_add_text` will be used to add text into that 
new paragraph.

```{r}
my_pres <- my_pres %>% 
  ph_add_par() %>%
  ph_add_text(str = "A small red text!", style = fp_text(color = "red", font.size = 10) ) 

my_pres <- my_pres %>% 
  ph_add_par(level = 2) %>%
  ph_add_text(str = "Level 2")
# print(my_pres, "my_pres.pptx") %>% browseURL()
```

### Add to an existing placeholder of text

The following code produce a presentation made of one text shape containing text "A first text".

```{r}
my_pres <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_text(type = "body", str = "A first text")
```

There is a paragraph in the new shape, function `ph_add_par` has to be used to add a new paragraph and `ph_add_text` has to be used to add text into the 
last paragraph of the shape.

```{r}
my_pres <- my_pres %>% 
  ph_add_text(str = "A small red text!", style = fp_text(color = "red", font.size = 10) ) %>% 
  ph_add_text(str = "Blue text first... ", pos = "before", style = fp_text(color = "blue", font.size = 0) )

# print(my_pres, "my_pres.pptx") %>% browseURL()
```


```{r echo=FALSE}
unlink("*.pptx")
```

