---
title: "Introduction to pptx"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pptx introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Creation of PowerPoint documents is made with function `pptx`. 

One can add and remove slides with functions `add_slide` and `remove_slide`. 
Function `on_slide` select a current slide. Management of the content of the 
current slide can be made with functions `placeholder_add_paragraph`, `placeholder_add_text`, 
`placeholder_remove`, `placeholder_set_img` and `placeholder_set_text`.


```{r}
library(officer)
library(magrittr)

doc <- pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme")
doc
```

```{r}
fileout <- tempfile(fileext = ".pptx")

doc %>%
  add_slide(layout = "Two Content", master = "Office Theme") %>%
  placeholder_set_text(id = "title", str = "A title") %>%
  placeholder_set_text(id = "ftr", str = "pied de page") %>%
  placeholder_set_text(id = "dt", str = format(Sys.Date())) %>%
  placeholder_set_text(str = "text 1", id = "body", index = 1) %>%
  placeholder_set_text(str = "text 2", id = "body", index = 2) %>%
  print(target = fileout )
browseURL(fileout)
```



