---
title: "docx device"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{docx device}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> `officer` provides a set of functions to manipulate Word documents from R.

*Package `magrittr` can be used with `officer` as it will make the workflow easier to read.*

```{r}
library(officer)
library(magrittr)
```

## Introduction

Use the function `read_docx` to create an r object representing a PowerPoint document.
It takes one argument `path`: a initial Word file. If none is provided, this file will be an empty document
located in the package directory. Formats and styles will be those available in the template file.

### Create a Word device

```{r}
my_doc <- read_docx() 
my_doc
```

### Add elements to document

Let's add an image in the documet first.

```{r}
src <- tempfile(fileext = ".png")
png(filename = src, width = 5, height = 6, units = 'in', res = 300)
barplot(1:10, col = 1:10)
dev.off()

my_doc <- my_doc %>% 
  body_add_img(src = src, width = 5, height = 6, style = "centered")
```

Then some paragraphs.

```{r}
my_doc <- my_doc %>% 
  body_add_par("Hello world!", style = "Normal") %>% 
  body_add_par("", style = "Normal") # blank paragraph
```

And a table.

```{r}
my_doc <- my_doc %>% 
  body_add_table(iris, style = "table_template", width = 3)
```

### Write the Word file

File can be generated using function `print` and argument `target`:

```{r}
outfile <- tempfile(fileext = ".docx")
print(my_doc, target = outfile )
```

## Getting started

To produce Word documents from R, you will need to manipulate few concepts.

`read_docx` read an initial `docx` file. It takes one argument `path`: a initial Word file (I sometimes refer to it as *template*. 
If none is provided, this file will be an empty document located in the package directory.
Styles will be those available in the template file.

### Add elements

To add paragraphs, tables, images or other elements into the document, you will have to use functions starting with `body_add_`:

```{r echo=FALSE}
objs <- objects(envir = as.environment("package:officer"), pattern = "^body_")
objs <- setdiff(objs, "body_add_xml")
htmltools::tags$ul(lapply(objs, htmltools::tags$li))
```

### Cursor

A cursor is available and can be manipulated so that content can be added regarding to its position: 

- `before` will insert a new element before the selected element in the document.
- `after` will insert a new element after the selected element in the document.
- `on` will replace the selected element in the document by a new element.

Cursor functions are the following:

```{r echo=FALSE}
objs <- objects(envir = as.environment("package:officer"), pattern = "^cursor_")
htmltools::tags$ul(lapply(objs, htmltools::tags$li))
```

### Append text in existing paragraphs

Text and images can be inserted at the beginning or the end of the selected paragraph (by the cursor). This can be made with functions 

```{r echo=FALSE}
objs <- objects(envir = as.environment("package:officer"), pattern = "^slip_in_")
objs <- setdiff(objs, "slip_in_xml")
htmltools::tags$ul(lapply(objs, htmltools::tags$li))
```


