---
title: "Officer and graphics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Officer and graphics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo = FALSE, message=FALSE, warning=FALSE}
dir.create("assets/graphics", recursive = TRUE, showWarnings = FALSE)
office_doc_link <- function(url){
  stopifnot(requireNamespace("htmltools", quietly = TRUE))
  htmltools::tags$p(  htmltools::tags$span("Download file "),
    htmltools::tags$a(basename(url), href = url), 
    htmltools::tags$span(" - view with"),
    htmltools::tags$a("office web viewer", target="_blank", 
      href = paste0("https://view.officeapps.live.com/op/view.aspx?src=", url)
      ), 
    style="text-align:center;font-style:italic;color:gray;"
    )
}
```

## Word documents and graphics 

Graphics and more generally raster images and vector graphic images are added in a Word document as elements of a **paragraph**. `officer` provides two functions to let users add images into a Word document:

- *body_add_img*: add an image in a new paragraph. This is usually the function to use when the need is to add a plot generated with R into a document.
- *slip_in_img*: add an image inside an existing paragraph (defined by the cursor location), user can choose to insert it at beginning or end of the paragraph. This allows to mix text and images within a single paragraph.

Those functions accept various image formats: png, jpeg and emf.

### R plots

So far, tests has been made with outputs from `grDevices::png()`, `grDevices::jpeg()` and `devEMF::emf()`. Enhanced metafile graphics produced by package `devEMF` seem to be the best solution to have vector graphics within Word document. 

```{r}
library(officer)
library(devEMF)
library(magrittr)
library(ggplot2)
```

`officer` is not producing the image, it only provides functions to add a produced image into a Word document.

Let's first produce the image:

```{r results='hide'}
# ggplot example
gg <- ggplot(mtcars, aes(x = mpg , y = wt, colour = qsec)) + geom_point() + theme_minimal()

# produce an emf file containing the ggplot
filename <- tempfile(fileext = ".emf")
emf(file = filename, width = 6, height = 7)
print(gg)
dev.off()
```

Then add it to a new Word document:

```{r}
read_docx() %>% 
  body_add_img(src = filename, width = 6, height = 7) %>% 
  print(target = "assets/graphics/demo_emf.docx")
```


```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/officer/articles/offcran/", "assets/graphics/demo_emf.docx" ) )
```

### Microsoft charts

Package `mschart` enable to create native office charts that can be used 
with `officer`. Function `body_add_chart` can be used to generate (from R plot
commands) a chart object within Word. 

```{r}
library(mschart)

my_barchart <- ms_barchart(data = browser_data,
  x = "browser", y = "value", group = "serie")
my_barchart <- chart_settings( x = my_barchart,
  dir="vertical", grouping="clustered", gap_width = 50 )

read_docx() %>% 
  body_add_chart(chart = my_barchart, style = "centered") %>% 
  print(target = "assets/graphics/demo_mschart_01.docx")
```



```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/officer/articles/offcran/", "assets/graphics/demo_mschart_01.docx" ) )
```

## PowerPoint documents and graphics 

Raster images and vector graphic images are added in a PowerPoint document as a 
shape on the current slide. In PowerPoint, images and text can not be mixed in 
the same shape (or placeholder). `officer` provides a function `ph_with` to let users add 
images into a PowerPoint document:

Again those functions accept various image formats: png, jpeg and emf.

```{r}
read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(external_img(src = filename, width = 6, height = 7), 
          location = ph_location_type(type = "body"), use_loc_size = FALSE ) %>% 
  add_slide(layout = "Title Only", master = "Office Theme") %>% 
  ph_with(external_img(src = filename, width = 6, height = 7), 
                 location = ph_location(left = 0, top = 0, width = 6, height = 7) ) %>% 
  print(target = "assets/graphics/demo_emf.pptx")
```



```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/officer/articles/offcran/", "assets/graphics/demo_emf.pptx" ) )
```

### vector graphics

Package `rvg` is providing editable vector graphics that can be used with
`officer`. Function `dml` can be used to generate (from R plot commands) a nice
editable graphic within PowerPoint. `ph_with` supports `dml` objects, you only
have to wrap your R graphic instructions (or ggplot object) with function `dml`.

```{r}
library(rvg)
read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(dml(ggobj = gg), location = ph_location_type(type = "body")) %>% 
  print(target = "assets/graphics/demo_rvg.pptx")
```



```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/officer/articles/offcran/", "assets/graphics/demo_rvg.pptx" ) )
```


### Microsoft charts

Package `mschart` enable to create native office charts that can be used 
with `officer`. Function `ph_with_chart` can be used to generate (from R plot
commands) a chart object within PowerPoint. Function `ph_with_chart_at` is also provided to let you add a chart at any arbitrary location in the current slide.

```{r}
library(mschart)

my_barchart <- ms_barchart(data = browser_data,
  x = "browser", y = "value", group = "serie")
my_barchart <- chart_settings( x = my_barchart,
  dir="vertical", grouping="clustered", gap_width = 50 )

read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_chart(chart = my_barchart, type = "body") %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_chart_at(chart = my_barchart, 
                   left = 0, top = 0, width = 4, height = 4) %>% 
  print(target = "assets/graphics/demo_mschart_01.pptx")
```



```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/officer/articles/offcran/", "assets/graphics/demo_mschart_01.pptx" ) )
```
