---
title: "PowerPoint presentations generation"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{PowerPoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
library(openssl)
library(htmltools)
opts_chunk$set(echo = TRUE)

add_table_print <- function(){
  table_print = function(x, ...) {
    if(nrow(x) > 20){
      x <- head(x, n = 20)
    }
    knitr::knit_print(knitr::kable(x))
  }
  # register the method
  registerS3method("knit_print", "data.frame", table_print)
  registerS3method("knit_print", "grouped_df", table_print)

}

image_html <- function(x, class = NULL){
  input <- normalizePath(x, mustWork = TRUE)
  base64 <- lapply(input, function(z){
    buf <- readBin(z, raw(), file.info(z)$size)
    base64_encode(buf, linebreaks = FALSE)
  })
  img <- lapply(base64, function(z){
    div( class = "img_slide", tags$img(src = paste0("data:image/png;base64,", z)) )
  })
  if( !is.null(class) )
    div(class=class, tagList(img))
  else div(tagList(img))
}
add_table_print()
pptx_to_image <- function( url, class = NULL ){
  file <- url
  cmd_ <- sprintf(
    "/Applications/LibreOffice.app/Contents/MacOS/soffice --headless --convert-to pdf --outdir %s %s",
    dirname(file), file )
  system(cmd_)
  
  pdf_file <- gsub("\\.(docx|pptx)$", ".pdf", file)
  screen_copies <- pdftools::pdf_convert(pdf = pdf_file, format = "png", verbose = FALSE)
  on.exit(unlink(c(pdf_file, file, screen_copies)))
  image_html(screen_copies, class = class)
}
```

```{r}
library(officer)
```


```{css echo=FALSE, eval=TRUE}
.contents > .row * > pre {
  font-size: 0.5em;
}
.img_slide img {
  border: gray 1px solid;
  width: 100%;
}
.row.vertical-align {
  display: flex;
  align-items: center;
}
```

The main functions for PowerPoint generatiion are presented below: 

![](https://www.ardata.fr/img/illustrations/officer_pptx.svg)


## Initialize

Use the function `read_pptx()` to create an R object representing a PowerPoint 
document. The initial PowerPoint file can be specified with the `path` argument. 
If none is provided, this file will be an empty document located in the package 
directory. Formats and available slide layouts will be those available in the 
template file. The content of original document is also preserved (but can be 
manipulated, i.e. delete a slide).
 
```{r}
my_pres <- read_pptx() 
```

## Add slides

To add a new slide, use the function `add_slide()`. It requires 3 arguments: 

- an rpptx object
- a slide layout name
- a master layout name

```{r}
# my_pres <- add_slide(my_pres, layout = "Title Only", master = "Office Theme")
# my_pres <- add_slide(my_pres, layout = "Two Content", master = "Office Theme")
my_pres <- add_slide(my_pres, layout = "Title and Content", master = "Office Theme")
```


<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

**Note that the `layout` and `master` values must match values from the initial document.** 

Layout names and master layout names are not easily *readable* within
PowerPoint, but these can be read easily with the function `layout_summary()`.

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r}
layout_summary(my_pres)
```

</div>

</div>

## Add content to a slide

Content must be placed on a slide in a new shape.

`ph_with()` is the unique function you should have to use to add content 
to a slide. It has 3 arguments, the `rpptx` object, the object to be 
printed and the location that define the placeholder where shape 
will be created. 

> Note that `ph_with` is to be used instead of 
functions `ph_with_*` and `ph_with_*_at`.


```{r}
my_pres <- ph_with(my_pres, value = "Hello world", location = ph_location_type(type = "title"))
my_pres <- ph_with(my_pres, value = "A footer", location = ph_location_type(type = "ftr"))
my_pres <- ph_with(my_pres, value = format(Sys.Date()), location = ph_location_type(type = "dt"))
my_pres <- ph_with(my_pres, value = "slide 1", location = ph_location_type(type = "sldNum"))
my_pres <- ph_with(my_pres, value = head(letters), location = ph_location_type(type = "body")) 
```

## Write the PowerPoint file

The (updated) Powerpoint file can be generated using the `print()` function 
along with the `target` argument:

```{r}
print(my_pres, target = "first_example.pptx") 
```

```{r echo=FALSE}
pptx_to_image(url = "first_example.pptx")
```


## Content location

The function `ph_with()` is expecting 
an argument `location` that will be used to specify the location of the new shape. 

This argument must be an object created from a call to one of the 
`ph_location*` functions (*placeholder location*). There are several 
`ph_location*` functions: `ph_location_type()`, `ph_location_fullsize()`,
`ph_location_label()`, `ph_location_left()`, `ph_location_right()`, 
`ph_location_template()` and `ph_location()`. 

----

<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

The following code adds text "Hello world" in a new shape. 

That shape will inherit its properties from the placeholder that has 
type `body`. This is defined in the slide layout used by the slide,
i.e. `Title and Content` in our example.

```{r}
doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(doc, c("Hello world", "Bonjour monde", "Blah blah blah"), 
          location = ph_location_type(type = "body") ) 

print(doc, target = "ph_with_location_type.pptx") 
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_location_type.pptx")
```

</div>

</div>

----

The following example will print few rows of `iris` data.frame in a new shape. That shape will 
inherit its properties from the placeholder that has label `Content Placeholder 2`. 

This is defined in the slide layout used by the slide, i.e. `Title and Content` 
in our example. Call `layout_properties()` to see available placeholder names (see 
column `ph_label`.

```{r}
mypres <- read_pptx()
mypres <- add_slide(mypres, layout = "Title and Content", master = "Office Theme") 
layout_properties ( x = mypres, layout = "Title and Content" )
```


<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

This is the same location that is used in the previous example 
but identified with function `ph_location_type()`. A title is 
also added.

```{r}
mypres <- ph_with(mypres, head(iris), 
  location = ph_location_label(
    ph_label = "Content Placeholder 2") )

mypres <- ph_with(mypres, "Slide title", 
          location = ph_location_label(
    ph_label = "Title 1") )

print(mypres, target = "ph_with_location.pptx") 
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_location.pptx")
```

</div>

</div>

----

<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

Ths example prints two paragraphs of text in two new shapes, 
one representing the left side and one representing the right side.

```{r}
mypres <- add_slide(mypres, layout = "Two Content")
mypres <- ph_with(mypres, sprintf("text item #%d", 1:6), 
          location = ph_location_left())
mypres <- ph_with(mypres, sprintf("text item #%d", 7:12), 
          location = ph_location_right()) 

print(mypres, target = "ph_with_location.pptx") 
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_location.pptx")
```

</div>

</div>


## Supported contents

The following types of content are supported: 

### Text and vectors

<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

```{r}
doc <- read_pptx()
doc <- add_slide(doc, layout = "Two Content", master = "Office Theme")
doc <- ph_with(doc, value = pi, location = ph_location_type(type = "title") )
doc <- ph_with(doc, value = as.factor(letters[1:2]), location = ph_location_type(type = "ftr") )
doc <- ph_with(doc, value = c("one blah", "two blah"), location = ph_location_left() )
doc <- ph_with(doc, value = c("three blah", "four blah"), location = ph_location_right() )
print(doc, target = "ph_with_vectors.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_vectors.pptx")
```

</div>

</div>

### data.frame

<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

```{r}
doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(x = doc, value = head(iris), 
   location = ph_location_type(type = "body") )
print(doc, target = "ph_with_df.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_df.pptx")
```

</div>

</div>

### ggplot objects


<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

```{r message=FALSE}
library(ggplot2)
gg_plot <- ggplot(data = iris ) +
  geom_point(
    mapping = aes(Sepal.Length, Petal.Length, color = Species),
    size = 3) + theme_minimal()

doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(x = doc, value = gg_plot, 
               location = ph_location_fullsize() )
doc <- ph_with(x = doc, "a ggplot example", 
               location = ph_location_type(
                 type = "title") )
print(doc, target = "ph_with_gg.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_gg.pptx")
```

</div>

</div>

### Images

> see wrapper function named `external_img()`

<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

```{r}
img.file <- file.path( R.home("doc"), "html", "logo.jpg" )

doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(x = doc, external_img(img.file, width = 1.39, height = 1.06),
   location = ph_location_type(type = "body"), use_loc_size = FALSE )
print(doc, target = "ph_with_img.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_img.pptx")
```

</div>

</div>


### Multiple paragraphs

<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">


> see wrapper function named `block_list()`

```{r}
# first define a block_list made of fpar ----
bl <- block_list(
  fpar(ftext("hello world", fp_text(bold = TRUE))),
  fpar(
    ftext("hello", fp_text(bold = TRUE, font.size = 30)),
    ftext(" world", prop = fp_text(color = "red", font.size = 30) )
  )
)

doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(x = doc, value = bl, 
               location = ph_location(label = "my_name",
                 left = 2, top = 3, width = 4, height = 4, 
                 bg = "wheat", rotation = 90) 
       )

print(doc, target = "ph_with_block_list.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_block_list.pptx")
```

</div>

</div>

### Single paragraph

> see wrapper function named `fpar()` (a `block_list` is 
 made of `fpar` objects)

<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

```{r}
# first define a fpar ----
fp <- fpar(
    ftext("hello", fp_text(bold = TRUE, font.size = 40)),
    ftext(" world", prop = fp_text(color = "red", font.size = 40) )
  )

doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(x = doc, value = fp, 
               location = ph_location_type(type = "title") 
       )

print(doc, target = "ph_with_fpar.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_fpar.pptx")
```

</div>

</div>

### Unordered lists

> See wrapper function named `unordered_list()`

<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

```{r}
doc <- read_pptx()

ul <- unordered_list(
  level_list = c(1, 2, 2, 3, 3, 1),
  str_list = c("Level1", "Level2", "Level2", "Level3", "Level3", "Level1"),
  style = fp_text(color = "red", font.size = 0) )

doc <- add_slide(doc)
doc <- ph_with(x = doc, value = ul, 
               location = ph_location_type(type = "body") )

print(doc, target = "ph_with_ul.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("ph_with_ul.pptx")
```

</div>

</div>

## Slide selection and manipulation

`add_slide()` is not the only function to let you manipulate slides:

* `remove_slide()`
* `move_slide()`
* `on_slide()`.

```{r}
my_pres <- read_pptx()
my_pres <- add_slide(my_pres, layout = "Title Only")
my_pres <- ph_with(my_pres, value = "Slide 1", location = ph_location_type(type = "title"))
my_pres <- add_slide(my_pres, layout = "Two Content")
my_pres <- ph_with(my_pres, value = "Slide 2", location = ph_location_type(type = "title"))
my_pres <- add_slide(my_pres, layout = "Title and Content")
my_pres <- ph_with(my_pres, value = "Slide 3", location = ph_location_type(type = "title"))
```



<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

A slide can be moved with the `move_slide()` function.

```{r}
my_pres <- move_slide(my_pres, index = 1, to = 3)
print(my_pres, target = "slide_manip.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("slide_manip.pptx")
```

</div>

</div>





<div class="row vertical-align">

<div class="col-12 col-sm-6 col-md-6">

A slide can be removed with the `remove_slide()` function.

```{r}
my_pres <- remove_slide(my_pres, index = 1)
print(my_pres, target = "slide_manip.pptx")
```

</div>

<div class="col-12 col-sm-6 col-md-6">

```{r echo=FALSE}
pptx_to_image("slide_manip.pptx")
```

</div>

</div>


A slide can be selected with the `on_slide()` function.

```{r}
my_pres <- on_slide(my_pres, index = 1)
```



## Remove content from a slide

Use `slide_summary()` to easily identify shapes in the slide that can be removed.

```{r}
slide_summary(doc)
```

In the following example, the shape corresponding to `type "body"` will be 
removed from the current slide:

```{r}
doc <- ph_remove(x = doc, type = "body", id = 1)
```


## Miscellaneous

### External links

`ph_hyperlink()` adds a hyperlink to an existing placeholder in the current 
slide. The argument `href` should contain a valid URL (i.e. starting with 
`http(s)`).

```{r}
doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(doc, "Blah blah blah", location = ph_location_type(type = "body"))
doc <- ph_hyperlink(doc, 
  ph_label = "Content Placeholder 2", href = "https://cran.r-project.org") 
```


### Internal links

`ph_slidelink()` adds an internal link into an existing placeholder. The 
argument `slide_index` should contain the index of the target slide.

```{r}
doc <- read_pptx()
doc <- add_slide(doc)
doc <- ph_with(doc, "Blah blah blah", location = ph_location_type(type = "body"))
doc <- add_slide(doc)
doc <- ph_with(doc, "placeholder target", location = ph_location_type(type = "title"))
doc <- on_slide(doc, index = 1 )
doc <- ph_slidelink(doc, 
  ph_label = "Content Placeholder 2", slide_index = 2)
```


## Resources

Len Kiefer wrote two very good blog posts about officer, he is providing nice examples with the corresponding R code:

* http://lenkiefer.com/2017/09/23/crafting-a-powerpoint-presentation-with-r/
* http://lenkiefer.com/2017/09/27/use-purrr/

[Eric Nanz](https://r-podcast.org) gave a short presentation about using officer in the Advanced R Markdown workshop held at the 2019 RStudio conference:

* https://rpodcast.github.io/officer-advrmarkdown/
